treats manual

### 2023-05-17 inst/gitbook/index.Rmd
Update index.Rmd

Ar Macarface is studying the evolution of mammalian disparity (i.e. diversity of shapes) across the Cretaceous-Palaeogene mass extinction event (K-Pg - 66 million years ago).
They observe a decrease of morphological disparity after the extinction event followed by a sharp increase of disparity.
However Ar is also prudent and wants to test whether their observed pattern is just an artifact of the data or actually showing some effect of the K-Pg extinction on mammalian evolution.
First they need to check whether changes in disparity can actually be discernables due to an extinction event (i.e. if an extinction event happens, would it be possible to detect a change in disparity)?
Second they want to know whether these changes are due to changes are due to the mass extinction event being selective or not.

The [getting started](#kpgexample) section actually shows how do implement this example in `treats` in details.

There are some very good R packages that simulate birth-death trees (e.g. [TreeSim](https://github.com/tanja819/TreeSim)) or simulate traits (disparity) and diversity jointly (e.g. [RPANDA](https://github.com/cran/RPANDA) or [PETER](https://github.com/PuttickMacroevolution/PETER)).
We strongly advise you look at these packages first as they might be more appropriate for your needs.

@@

![Schematic summary of the `treats` package architecture](treats_structure.png)
% NC: The wonkiness of the figure is annoying me lol. Can you center the arrows etc?

>>>>

Bib is studying the evolution of mammalian disparity (i.e. diversity of shapes) across the Cretaceous-Palaeogene mass extinction event (K-Pg - 66 million years ago).
They observe a decrease of morphological disparity after the extinction event followed by a sharp increase of disparity.
However Bib is also prudent and wants to test whether their observed pattern is just an artifact of the data or actually showing some effect of the K-Pg extinction on mammalian evolution.
First they need to check whether changes in disparity can actually be discernables due to an extinction event (i.e. if an extinction event happens, would it be possible to detect a change in disparity)?
Second they want to know whether these changes are due to the mass extinction event being selective or not (i.e. the mass extinction favored some traits).

The [illustrative example](#kpgexample) section actually shows how do implement this example in `treats` in details.

Of course, this is just an example of when to use `treats`.
Simulating trees and traits can also be useful for:

 * Provide a baseline/null model to compare to an observed one.
 * Testing evolutionary hypotheses.
 * Exploring statistics properties (i.e. can the disparity metric X capture changes in Y).
 * Easily generating illustrative/simplified evolutionary history examples.
 * Teaching simplified macro-evolutionary theory (e.g. the birth-death model).

Note that for approaching these points individually (or some aspects of it), there are some very good R packages that simulate birth-death trees (e.g. [TreeSim](https://github.com/tanja819/TreeSim)) or simulate traits (disparity) and diversity jointly (e.g. [RPANDA](https://github.com/cran/RPANDA) or [PETER](https://github.com/PuttickMacroevolution/PETER)).
We strongly advise you look at these packages first as they might be more appropriate for your needs.

@@

![Schematic summary of the `treats` package architecture](treats_structure.png)
